\documentclass{article}
\usepackage{enumitem}
\usepackage[margin=3cm]{geometry}
\title{Bachelor Meeting Endri Dibra}
\author{Joanna Brodbeck}
\date{\today}

\begin{document}
\maketitle


\subsubsection*{Skills}
\begin{itemize}
    \item Machine Learning
    \begin{itemize}
        \item How to predict with missing data?
        \item How to use least-squares
        \item KFold with different lambdas
        \item How to work with images in the machine learning part and neural networks context
    \end{itemize}
    \item Computer Vision
    \begin{itemize}
        \item since September visiting Visual Computing
        \item Next semester: Visualization, Simulation and Interaction I
        \item Ready to catch up the other missing pieces
    \end{itemize}
    \item Programming
    \begin{itemize}
        \item Reliable Transport Protocol (C)
        \item Machine Learning Projects (Python)
        \item Wooden Box (Python)
        \item Whatsapp Bot (Python) (current)
        \item Countdown App (Android Studios) (current)
        \item Website (Typescript) (current)
    \end{itemize}
\end{itemize}

\subsubsection*{Motivation}
\begin{itemize}
    \item Medical aspect from family: Sister studied Medicine
    \item Cancer case in my family
    \item Want to make the world a better place
    \item Love to computer science
    \item Programming is a hobby of mine, which I am passionate about.
    \item Love to learn new things
    \item I wonder how different kind of things work.
\end{itemize}

\subsubsection*{Questions}

\begin{enumerate}[label=\textbf{\arabic*.}]
    \item What if I am stuck in the process of the bachelor thesis?
    \item What language am I programming in?
    \item What is the deadline for the bachelor thesis?
    \item How many meetings are there? (1x per week?)
    \item Who am I working with? Alone? Team?
    \item Explain further what the thesis are about?
    \item Where would I work, in the breast or face department?
    \item Is it a project that the app is open for patients / general users
    \item How much work is expected?
    \item Duration and when to start?
    \item Communication other than Email?
    \item What kind of guidance can I expected?
    \item Things to learn before the start of the thesis?







\end{enumerate}

\end{document}
